---
title: "BA1 geophysical/chemical logs"
subtitle: "Source file: 201012_BA1_Logs.Rmd"
author: "Daniel Nothaft"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged # omit to disable paged data table output
    css: stylesheet.css # omit if no need for custom stylesheet
    number_sections: yes # change to no for unnumbered sections
    toc: yes # change to no to disable table of contents
    toc_float: true # change to false to keep toc at the top
    toc_depth: 3 # change to specify which headings to include in toc
    code_folding: show # change to hide to hide code by default
editor_options:
  chunk_output_type: inline
---
# Setup
## Load code libraries
```{r setup, message=FALSE, warning=FALSE}
# load libraries
library(tidyverse) # dplyr, tidyr, ggplot
library(readxl) # reading excel files
library(knitr) # generating reports
library(ggpmisc) # add equations of best fit to ggplot
library(scales) # making log scales with exponents
library(lemon) # repeat axis lines for paneled figures in ggplot

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"), fig.keep = "all",
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.path = file.path("fig_output/", paste0(gsub("\\.[Rr]md", "/", knitr::current_input()))),
  cache.path = file.path("cache/", paste0(gsub("\\.[Rr]md", "/", knitr::current_input())))
)
```

```{r source}
# source all relevant scripting files
source(file.path("scripts", "plotting_functions.R"))
```

## Load data

```{r load-data}
# read BA1A borehole log from March 16th, 2018
BA1A_2018_profile_raw <- read_xlsx("data_raw/Summary_metadata/BA1A_QL40IDRO_20180316_DOWN_5cm.xlsx", na = c("-99999.00", "-99999"))

# # read BA1A borehole log from April 22nd, 2017
# BA1A_2017_profile_raw <- read_xlsx("data_raw/Summary_metadata/BA1A_Muset_22APRIL2017_run2_pass1_dw_5cm.xlsx", na = c("-99999.00", "-99999"))

# # read BA1A borehole log from March 19th, 2018
# BA1D_2018_profile_raw <- read_xlsx("data_raw/Summary_metadata/BA1D_QL40IDRO_20180319_down_5cm.xlsx", na = c("-99999.00", "-99999"))

# read BA1A borehole log from March 19th, 2018
BA1D_2019_profile_raw <- read_xlsx("data_raw/Summary_metadata/BA1D_2019_chemical_log.xlsx", na = c("-99999.00", "-99999"))

# For context, well BA1A was drilled in 2017, from February 20 to March 2, and BA1D was drilled in 2018, from February 24 to March 15.

# read Oman Geochem data, including data from packer sampling
Oman_Geochem <- read_xlsx("data_raw/Summary_metadata/Oman_Geochem_2012-2019_packers.xlsx", na = c("NA"))

# Read in conversion from Ag/AgCl electrode to SHE as function of T
Ag_AgCl_E_T <- read_xlsx("data_raw/Summary_metadata/Ag_AgCl_E_T.xlsx", na = "NA")
```

# BA1A 2018 well log

## clean up data
```{r get-relevant-log-data-BA1A-2018}
# select relevant columns and rename them
BA1A_2018_profile <- BA1A_2018_profile_raw %>% select(depth_sampled_mbgl = depth_m, P_dbar = pressure, T_C = Temp_C, pH, ORP_mV_Ag_AgCl = ORP, cond_uS_per_cm = Cond_FW)

# Filter out nonsense values (i.e. when probe is above water table)
BA1A_2018_profile <- BA1A_2018_profile %>% filter_all(all_vars(!is.na(.)))

# convert conductivity to mS/cm
BA1A_2018_profile <- BA1A_2018_profile %>% mutate(cond_mS_per_cm = cond_uS_per_cm * 1e-3) %>% select(-cond_uS_per_cm)

# print some summary statistics
BA1A_2018_profile %>% summary()

BA1A_2018_profile %>% head()
```

```{r subset-and-print-log-BA1A_2018}
# subset the data for readability and print
BA1A_2018_profile %>% slice(seq(from = 1, to = 10000, by = 200)) %>% kable(digits=1)
```

## Convert to ORP to SHE

Plot conversion between Ag/AgCl electrode and SHE as function of temperature
```{r plot-Ag-AgCl-SHE-conversion}
formula <- Ag_AgCl_E_T$E_Ag_AgCl_KCl_sat_mV ~ Ag_AgCl_E_T$T_C

plot_redox_offset_T <- Ag_AgCl_E_T %>% ggplot(aes(x = T_C, y = E_Ag_AgCl_KCl_sat_mV))+
  geom_point()+
  geom_smooth(method="lm")+
  stat_poly_eq(aes(label =  paste(stat(eq.label), stat(rr.label), sep = "~~~~")),
               formula = formula, parse = TRUE, rr.digits = 2, label.x = .2, label.y = .9)+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank())

plot_redox_offset_T
```
Save a linear model of the relationship between Ag/AgCl electrode and SHE as function of temperature
```{r save-lm-Ag-AgCl-SHE}
Ag_AgCl_E_T_lm <- lm(formula)

Ag_AgCl_E_T_lm_summ <- Ag_AgCl_E_T_lm %>% summary()

Ag_AgCl_E_T_lm_summ
```
Save slope
```{r save-lm-Ag-AgCl-SHE-slope}
Ag_AgCl_E_T_lm_slope <- coef(Ag_AgCl_E_T_lm_summ)["Ag_AgCl_E_T$T_C",  "Estimate"]

Ag_AgCl_E_T_lm_slope
```
Save intercept
```{r save-lm-Ag-AgCl-SHE-int}
Ag_AgCl_E_T_lm_int <- coef(Ag_AgCl_E_T_lm_summ)["(Intercept)",  "Estimate"]

Ag_AgCl_E_T_lm_int
```

Apply conversion
```{r apply-Ag-AgCl-SHE-conversion-NSHQ14}
BA1A_2018_profile <- BA1A_2018_profile %>% mutate(Eh_mV = ORP_mV_Ag_AgCl + Ag_AgCl_E_T_lm_int + Ag_AgCl_E_T_lm_slope * T_C) # convert probe ORP_mV_Ag_AgCl reading from Ag/AgCl to SHE

# print some summary statistics
BA1A_2018_profile %>% select(Eh_mV) %>% summary()
```

## Hydrostatic pressure
Pressure will be used to assess calculated stability limit of water with respect to molecular hydrogen.

Calculate from theory
```{r calc-hydrostatic-P-NSHQ14}
P_BA1A_2018_atmospheric_Pa <- 95490 # measured ambient pressure with Garmin GPSMAP 64S by DBN in 2019
P_BA1A_2018_atmospheric_bar <-  P_BA1A_2018_atmospheric_Pa * 1e-5

water_level_BA1A_2018_m <- as.numeric(BA1A_2018_profile %>% select(depth_sampled_mbgl) %>% min())

BA1A_2018_profile <- BA1A_2018_profile %>% mutate(water_column_above_m = depth_sampled_mbgl - water_level_BA1A_2018_m)


# Hydrostatic pressure in a liquid can be calculated as
# 
# p = ρ g h                         (1)
# 
# where
# 
# p = pressure in liquid (N/m2, Pa, lbf/ft2, psf)
# 
# ρ = density of liquid (kg/m3, slugs/ft3)
# 
# g = acceleration of gravity (9.81 m/s2, 32.17405 ft/s2)
# 
# h = height of fluid column - or depth in the fluid where pressure is measured (m, ft)

# https://www.engineeringtoolbox.com/hydrostatic-pressure-water-d_1632.html

density_water <- 1000 # kg/m3
g = 9.81 # acceleration of gravity, m/s2


BA1A_2018_profile <- BA1A_2018_profile %>% mutate(P_hydrostatic_calc_Pa = P_BA1A_2018_atmospheric_Pa + density_water * g * water_column_above_m)

BA1A_2018_profile <- BA1A_2018_profile %>% mutate(P_hydrostatic_calc_bar = P_hydrostatic_calc_Pa * 1e-5)

# print some summary statistics
BA1A_2018_profile %>% select(P_hydrostatic_calc_bar) %>% summary()
```

```{r convert-hydrostatic-P-to-bar-BA1A_2018}
# convert measured pressure to bar
BA1A_2018_profile <- BA1A_2018_profile %>% mutate(P_bar = P_dbar / 10)
```

```{r prepare-P-plot-BA1A_2018}
# prepare for plotting
BA1A_2018_profile_longer_P <- BA1A_2018_profile %>% select(depth_sampled_mbgl, P_hydrostatic_calc_bar, P_bar) %>% gather(-depth_sampled_mbgl, key = parameter, value = P_bar) %>% mutate(type = if_else(parameter == "P_hydrostatic_calc_bar", "calculated", "measured"))
```

Compare theoretical hydrostatic pressure versus measured pressure
```{r P-plot-BA1A_2018, fig.height=4, fig.width=4}
plot_BA1A_2018_P <- BA1A_2018_profile_longer_P %>% ggplot(aes(x = depth_sampled_mbgl))+
  geom_line(aes(y = P_bar, color = type))+
  scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02), breaks = c(0, 50, 100, 150, 200, 250, 300))+
    scale_y_continuous(name = latex2exp::TeX("$\\textit{P}_{hydrostatic}\\,/\\,\\lbrack bar \\rbrack$"))+
  coord_flip()+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank())

plot_BA1A_2018_P
```
They are similar.

## Stability limits of water

Eh, f(T,P)
```{r calc-H2-H20-limit-BA1A_2018}
Faraday <- 96485.33212331 # C/mol
R <- 8.31446261815324 # J/(mol⋅K) gas constant

BA1A_2018_profile <- BA1A_2018_profile %>% mutate(T_K= T_C + 273.15) # convert T to Kelvins

# calculate H20-H2 stability limit
BA1A_2018_profile <- BA1A_2018_profile %>% mutate(H2O_H2_limit_Eh_f_T_P_mV = (-R*T_K/2/Faraday * log(((10^-pH)^-2)*P_bar))*1000)

BA1A_2018_profile %>% select(H2O_H2_limit_Eh_f_T_P_mV) %>% summary()
```

```{r prepare-H2-H20-limit-plot-BA1A_2018}
# prepare for plotting
BA1A_2018_profile_longer_Eh_T_P <- BA1A_2018_profile %>% select(depth_sampled_mbgl, Eh_mV, H2O_H2_limit_Eh_f_T_P_mV) %>% gather(-depth_sampled_mbgl, key = parameter, value = Eh_mV)
```

```{r H2-H20-limit-plot-BA1A_2018, fig.height=5, fig.width=4}
plot_BA1A_2018_Eh <- BA1A_2018_profile_longer_Eh_T_P %>% ggplot(aes(x = depth_sampled_mbgl))+
  geom_line(aes(y = Eh_mV, color = parameter))+
  scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02))+
  scale_y_continuous(name = "Eh / [mV]")+
  coord_flip()+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "vertical")

plot_BA1A_2018_Eh
```

# BA1A 2017 well log

## clean up data
```{r get-relevant-log-data-BA1A_2017}
# # select relevant columns and rename them
# BA1A_2017_profile <- BA1A_2017_profile_raw %>% select(depth_sampled_mbgl = `DEPT[M]`, P_dbar = PRESSURE, T_C = TEMP, pH = PH, ORP_mV_Ag_AgCl = EH, cond_mS_per_cm = COND)
# 
# # Filter out nonsense values (i.e. when probe is above water table)
# BA1A_2017_profile <- BA1A_2017_profile %>% filter_all(all_vars(!is.na(.)))
# 
# # print some summary statistics
# BA1A_2017_profile %>% summary()
```

```{r subset-and-print-log-BA1A_2017}
# # subset the data for readability and print
# BA1A_2017_profile %>% slice(seq(from = 1, to = 40000, by = 1000)) %>% kable(digits=1)
```

## Convert to ORP to SHE

Apply conversion
```{r apply-Ag-AgCl-SHE-conversion-BA1A_2017}
# BA1A_2017_profile <- BA1A_2017_profile %>% mutate(Eh_mV = ORP_mV_Ag_AgCl + Ag_AgCl_E_T_lm_int + Ag_AgCl_E_T_lm_slope * T_C) # convert probe ORP_mV_Ag_AgCl reading from Ag/AgCl to SHE
# 
# # print some summary statistics
# BA1A_2017_profile %>% select(Eh_mV) %>% summary()
```

## Hydrostatic pressure
Pressure will be used to assess calculated stability limit of water with respect to molecular hydrogen.

Calculate from theory
```{r calc-hydrostatic-P-BA1A_2017}
# P_BA1A_2017_atmospheric_Pa <- 95490 # measured ambient pressure with Garmin GPSMAP 64S by DBN in 2019
# P_BA1A_2017_atmospheric_bar <-  P_BA1A_2017_atmospheric_Pa * 1e-5
# 
# water_level_BA1A_2017_m <- as.numeric(BA1A_2017_profile %>% select(depth_sampled_mbgl) %>% min())
# 
# BA1A_2017_profile <- BA1A_2017_profile %>% mutate(water_column_above_m = depth_sampled_mbgl - water_level_BA1A_2017_m)
# 
# 
# # Hydrostatic pressure in a liquid can be calculated as
# # 
# # p = ρ g h                         (1)
# # 
# # where
# # 
# # p = pressure in liquid (N/m2, Pa, lbf/ft2, psf)
# # 
# # ρ = density of liquid (kg/m3, slugs/ft3)
# # 
# # g = acceleration of gravity (9.81 m/s2, 32.17405 ft/s2)
# # 
# # h = height of fluid column - or depth in the fluid where pressure is measured (m, ft)
# 
# # https://www.engineeringtoolbox.com/hydrostatic-pressure-water-d_1632.html
# 
# density_water <- 1000 # kg/m3
# g = 9.81 # acceleration of gravity, m/s2
# 
# 
# BA1A_2017_profile <- BA1A_2017_profile %>% mutate(P_hydrostatic_calc_Pa = P_BA1A_2017_atmospheric_Pa + density_water * g * water_column_above_m)
# 
# BA1A_2017_profile <- BA1A_2017_profile %>% mutate(P_hydrostatic_calc_bar = P_hydrostatic_calc_Pa * 1e-5)
# 
# # print some summary statistics
# BA1A_2017_profile %>% select(P_hydrostatic_calc_bar) %>% summary()
```
```{r convert-hydrostatic-P-to-bar-BA1A_2017}
# # convert measured pressure to bar
# BA1A_2017_profile <- BA1A_2017_profile %>% mutate(P_bar = P_dbar / 10)
```

```{r prepare-P-plot-BA1A_2017}
# # prepare for plotting
# BA1A_2017_profile_longer_P <- BA1A_2017_profile %>% select(depth_sampled_mbgl, P_hydrostatic_calc_bar, P_bar) %>% gather(-depth_sampled_mbgl, key = parameter, value = P_bar) %>% mutate(type = if_else(parameter == "P_hydrostatic_calc_bar", "calculated", "measured"))
```

Compare theoretical hydrostatic pressure versus measured pressure
```{r P-plot-BA1A_2017, fig.height=4, fig.width=4}
# plot_BA1A_2017_P <- BA1A_2017_profile_longer_P %>% ggplot(aes(x = depth_sampled_mbgl))+
#   geom_line(aes(y = P_bar, color = type))+
#   scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02), breaks = c(0, 50, 100, 150, 200, 250, 300))+
#     scale_y_continuous(name = latex2exp::TeX("$\\textit{P}_{hydrostatic}\\,/\\,\\lbrack bar \\rbrack$"))+
#   coord_flip()+
#   theme_bw(base_size = 12)+
#   theme(
#     panel.grid = element_blank())
# 
# plot_BA1A_2017_P
```
They are similar.

## Stability limits of water

Eh, f(T,P)
```{r calc-H2-H20-limit-BA1A_2017}
# Faraday <- 96485.33212331 # C/mol
# R <- 8.31446261815324 # J/(mol⋅K) gas constant
# 
# BA1A_2017_profile <- BA1A_2017_profile %>% mutate(T_K= T_C + 273.15) # convert T to Kelvins
# 
# # calculate H20-H2 stability limit
# BA1A_2017_profile <- BA1A_2017_profile %>% mutate(H2O_H2_limit_Eh_f_T_P_mV = (-R*T_K/2/Faraday * log(((10^-pH)^-2)*P_bar))*1000)
# 
# BA1A_2017_profile %>% select(H2O_H2_limit_Eh_f_T_P_mV) %>% summary()
```

```{r prepare-H2-H20-limit-plot-BA1A_2017}
# # prepare for plotting
# BA1A_2017_profile_longer_Eh_T_P <- BA1A_2017_profile %>% select(depth_sampled_mbgl, Eh_mV, H2O_H2_limit_Eh_f_T_P_mV) %>% gather(-depth_sampled_mbgl, key = parameter, value = Eh_mV)
```

```{r H2-H20-limit-plot-BA1A_2017, fig.height=5, fig.width=4}
# plot_BA1A_2017_Eh <- BA1A_2017_profile_longer_Eh_T_P %>% ggplot(aes(x = depth_sampled_mbgl))+
#   geom_line(aes(y = Eh_mV, color = parameter))+
#   scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02))+
#   scale_y_continuous(name = "Eh / [mV]")+
#   coord_flip()+
#   theme_bw(base_size = 12)+
#   theme(
#     panel.grid = element_blank(),
#     legend.position = "bottom",
#     legend.direction = "vertical")
# 
# plot_BA1A_2017_Eh
```

# BA1D 2018 well log

## clean up data
```{r get-relevant-log-data-BA1D_2018}
# # select relevant columns and rename them
# BA1D_2018_profile <- BA1D_2018_profile_raw %>% select(depth_sampled_mbgl = `DEPT[M]`, P_dbar = PRESSURE, T_C = TEMPERATURE, pH = PH, ORP_mV_Ag_AgCl = REDOX, cond_uS_per_cm = `COND(FW)`)
# 
# 
# # Filter out nonsense values (i.e. when probe is above water table)
# BA1D_2018_profile <- BA1D_2018_profile %>% filter_all(all_vars(!is.na(.)))
# 
# # convert conductivity to mS/cm
# BA1D_2018_profile <- BA1D_2018_profile %>% mutate(cond_mS_per_cm = cond_uS_per_cm * 1e-3) %>% select(-cond_uS_per_cm)
# # 
# # # print some summary statistics
# BA1D_2018_profile %>% summary()
```

```{r subset-and-print-log-BA1D_2018}
# # subset the data for readability and print
# BA1D_2018_profile %>% slice(seq(from = 1, to = 40000, by = 1000)) %>% kable(digits=1)
```

## Convert to ORP to SHE

Apply conversion
```{r apply-Ag-AgCl-SHE-conversion-BA1D_2018}
# BA1D_2018_profile <- BA1D_2018_profile %>% mutate(Eh_mV = ORP_mV_Ag_AgCl + Ag_AgCl_E_T_lm_int + Ag_AgCl_E_T_lm_slope * T_C) # convert probe ORP_mV_Ag_AgCl reading from Ag/AgCl to SHE
# 
# # print some summary statistics
# BA1D_2018_profile %>% select(Eh_mV) %>% summary()
```

## Hydrostatic pressure
Pressure will be used to assess calculated stability limit of water with respect to molecular hydrogen.

Calculate from theory
```{r calc-hydrostatic-P-BA1D_2018}
# P_BA1D_2018_atmospheric_Pa <- 95490 # measured ambient pressure with Garmin GPSMAP 64S by DBN in 2019
# P_BA1D_2018_atmospheric_bar <-  P_BA1D_2018_atmospheric_Pa * 1e-5
# 
# water_level_BA1D_2018_m <- as.numeric(BA1D_2018_profile %>% select(depth_sampled_mbgl) %>% min())
# 
# BA1D_2018_profile <- BA1D_2018_profile %>% mutate(water_column_above_m = depth_sampled_mbgl - water_level_BA1D_2018_m)
# 
# 
# # Hydrostatic pressure in a liquid can be calculated as
# # 
# # p = ρ g h                         (1)
# # 
# # where
# # 
# # p = pressure in liquid (N/m2, Pa, lbf/ft2, psf)
# # 
# # ρ = density of liquid (kg/m3, slugs/ft3)
# # 
# # g = acceleration of gravity (9.81 m/s2, 32.17405 ft/s2)
# # 
# # h = height of fluid column - or depth in the fluid where pressure is measured (m, ft)
# 
# # https://www.engineeringtoolbox.com/hydrostatic-pressure-water-d_1632.html
# 
# density_water <- 1000 # kg/m3
# g = 9.81 # acceleration of gravity, m/s2
# 
# 
# BA1D_2018_profile <- BA1D_2018_profile %>% mutate(P_hydrostatic_calc_Pa = P_BA1D_2018_atmospheric_Pa + density_water * g * water_column_above_m)
# 
# BA1D_2018_profile <- BA1D_2018_profile %>% mutate(P_hydrostatic_calc_bar = P_hydrostatic_calc_Pa * 1e-5)
# 
# # print some summary statistics
# BA1D_2018_profile %>% select(P_hydrostatic_calc_bar) %>% summary()
```
```{r convert-hydrostatic-P-to-bar-BA1D_2018}
# # convert measured pressure to bar
# BA1D_2018_profile <- BA1D_2018_profile %>% mutate(P_bar = P_dbar / 10)
```

```{r prepare-P-plot-BA1D_2018}
# # prepare for plotting
# BA1D_2018_profile_longer_P <- BA1D_2018_profile %>% select(depth_sampled_mbgl, P_hydrostatic_calc_bar, P_bar) %>% gather(-depth_sampled_mbgl, key = parameter, value = P_bar) %>% mutate(type = if_else(parameter == "P_hydrostatic_calc_bar", "calculated", "measured"))
```

Compare theoretical hydrostatic pressure versus measured pressure
```{r P-plot-BA1D_2018, fig.height=4, fig.width=4}
# plot_BA1D_2018_P <- BA1D_2018_profile_longer_P %>% ggplot(aes(x = depth_sampled_mbgl))+
#   geom_line(aes(y = P_bar, color = type))+
#   scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02), breaks = c(0, 50, 100, 150, 200, 250, 300))+
#     scale_y_continuous(name = latex2exp::TeX("$\\textit{P}_{hydrostatic}\\,/\\,\\lbrack bar \\rbrack$"))+
#   coord_flip()+
#   theme_bw(base_size = 12)+
#   theme(
#     panel.grid = element_blank())
# 
# plot_BA1D_2018_P
```
They are similar.

## Stability limits of water

Eh, f(T,P)
```{r calc-H2-H20-limit-BA1D_2018}
# Faraday <- 96485.33212331 # C/mol
# R <- 8.31446261815324 # J/(mol⋅K) gas constant
# 
# BA1D_2018_profile <- BA1D_2018_profile %>% mutate(T_K= T_C + 273.15) # convert T to Kelvins
# 
# # calculate H20-H2 stability limit
# BA1D_2018_profile <- BA1D_2018_profile %>% mutate(H2O_H2_limit_Eh_f_T_P_mV = (-R*T_K/2/Faraday * log(((10^-pH)^-2)*P_bar))*1000)
# 
# BA1D_2018_profile %>% select(H2O_H2_limit_Eh_f_T_P_mV) %>% summary()
```

```{r prepare-H2-H20-limit-plot-BA1D_2018}
# # prepare for plotting
# BA1D_2018_profile_longer_Eh_T_P <- BA1D_2018_profile %>% select(depth_sampled_mbgl, Eh_mV, H2O_H2_limit_Eh_f_T_P_mV) %>% gather(-depth_sampled_mbgl, key = parameter, value = Eh_mV)
```

```{r H2-H20-limit-plot-BA1D_2018, fig.height=5, fig.width=4}
# plot_BA1D_2018_Eh <- BA1D_2018_profile_longer_Eh_T_P %>% ggplot(aes(x = depth_sampled_mbgl))+
#   geom_line(aes(y = Eh_mV, color = parameter))+
#   scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02))+
#   scale_y_continuous(name = "Eh / [mV]")+
#   coord_flip()+
#   theme_bw(base_size = 12)+
#   theme(
#     panel.grid = element_blank(),
#     legend.position = "bottom",
#     legend.direction = "vertical")
# 
# plot_BA1D_2018_Eh
```

# BA1D 2019 well log

## clean up data
```{r get-relevant-log-data-BA1D_2019}
# select relevant columns and rename them
BA1D_2019_profile <- BA1D_2019_profile_raw %>% select(depth_sampled_mbgl = `Depth [m]`, P_dbar = `Pressure [dbar]`, T_C = `Temperature [degC]`, pH = pH, ORP_mV_Ag_AgCl = `Redox Potential [mV]`, cond_uS_per_cm = `Conductivity [uS/cm]`)

# Filter out nonsense values (i.e. when probe is above water table)
BA1D_2019_profile <- BA1D_2019_profile %>% filter_all(all_vars(!is.na(.)))

# convert conductivity to mS/cm
BA1D_2019_profile <- BA1D_2019_profile %>% mutate(cond_mS_per_cm = cond_uS_per_cm * 1e-3) %>% select(-cond_uS_per_cm)
# 
# # print some summary statistics
BA1D_2019_profile %>% summary()
```

```{r subset-and-print-log-BA1D_2019}
# subset the data for readability and print
BA1D_2019_profile %>% slice(seq(from = 1, to = 40000, by = 1000)) %>% kable(digits=1)
```

## Convert to ORP to SHE

Apply conversion
```{r apply-Ag-AgCl-SHE-conversion-BA1D_2019}
BA1D_2019_profile <- BA1D_2019_profile %>% mutate(Eh_mV = ORP_mV_Ag_AgCl + Ag_AgCl_E_T_lm_int + Ag_AgCl_E_T_lm_slope * T_C) # convert probe ORP_mV_Ag_AgCl reading from Ag/AgCl to SHE

# print some summary statistics
BA1D_2019_profile %>% select(Eh_mV) %>% summary()
```

## Hydrostatic pressure
Pressure will be used to assess calculated stability limit of water with respect to molecular hydrogen.

Calculate from theory
```{r calc-hydrostatic-P-BA1D_2019}
P_BA1D_2019_atmospheric_Pa <- 95490 # measured ambient pressure with Garmin GPSMAP 64S by DBN in 2019
P_BA1D_2019_atmospheric_bar <-  P_BA1D_2019_atmospheric_Pa * 1e-5

water_level_BA1D_2019_m <- as.numeric(BA1D_2019_profile %>% select(depth_sampled_mbgl) %>% min())

BA1D_2019_profile <- BA1D_2019_profile %>% mutate(water_column_above_m = depth_sampled_mbgl - water_level_BA1D_2019_m)


# Hydrostatic pressure in a liquid can be calculated as
# 
# p = ρ g h                         (1)
# 
# where
# 
# p = pressure in liquid (N/m2, Pa, lbf/ft2, psf)
# 
# ρ = density of liquid (kg/m3, slugs/ft3)
# 
# g = acceleration of gravity (9.81 m/s2, 32.17405 ft/s2)
# 
# h = height of fluid column - or depth in the fluid where pressure is measured (m, ft)

# https://www.engineeringtoolbox.com/hydrostatic-pressure-water-d_1632.html

density_water <- 1000 # kg/m3
g = 9.81 # acceleration of gravity, m/s2


BA1D_2019_profile <- BA1D_2019_profile %>% mutate(P_hydrostatic_calc_Pa = P_BA1D_2019_atmospheric_Pa + density_water * g * water_column_above_m)

BA1D_2019_profile <- BA1D_2019_profile %>% mutate(P_hydrostatic_calc_bar = P_hydrostatic_calc_Pa * 1e-5)

# print some summary statistics
BA1D_2019_profile %>% select(P_hydrostatic_calc_bar) %>% summary()
```
```{r convert-hydrostatic-P-to-bar-BA1D_2019}
# convert measured pressure to bar
BA1D_2019_profile <- BA1D_2019_profile %>% mutate(P_bar = P_dbar / 10)
```

```{r prepare-P-plot-BA1D_2019}
# prepare for plotting
BA1D_2019_profile_longer_P <- BA1D_2019_profile %>% select(depth_sampled_mbgl, P_hydrostatic_calc_bar, P_bar) %>% gather(-depth_sampled_mbgl, key = parameter, value = P_bar) %>% mutate(type = if_else(parameter == "P_hydrostatic_calc_bar", "calculated", "measured"))
```

Compare theoretical hydrostatic pressure versus measured pressure
```{r P-plot-BA1D_2019, fig.height=4, fig.width=4}
plot_BA1D_2019_P <- BA1D_2019_profile_longer_P %>% ggplot(aes(x = depth_sampled_mbgl))+
  geom_line(aes(y = P_bar, color = type))+
  scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02))+
    scale_y_continuous(name = latex2exp::TeX("$\\textit{P}_{hydrostatic}\\,/\\,\\lbrack bar \\rbrack$"))+
  coord_flip()+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank())

plot_BA1D_2019_P
```
They are similar.

## Stability limits of water

Eh, f(T,P)
```{r calc-H2-H20-limit-BA1D_2019}
Faraday <- 96485.33212331 # C/mol
R <- 8.31446261815324 # J/(mol⋅K) gas constant

BA1D_2019_profile <- BA1D_2019_profile %>% mutate(T_K= T_C + 273.15) # convert T to Kelvins

# calculate H20-H2 stability limit
BA1D_2019_profile <- BA1D_2019_profile %>% mutate(H2O_H2_limit_Eh_f_T_P_mV = (-R*T_K/2/Faraday * log(((10^-pH)^-2)*P_bar))*1000)

BA1D_2019_profile %>% select(H2O_H2_limit_Eh_f_T_P_mV) %>% summary()
```

```{r prepare-H2-H20-limit-plot-BA1D_2019}
# prepare for plotting
BA1D_2019_profile_longer_Eh_T_P <- BA1D_2019_profile %>% select(depth_sampled_mbgl, Eh_mV, H2O_H2_limit_Eh_f_T_P_mV) %>% gather(-depth_sampled_mbgl, key = parameter, value = Eh_mV)
```

```{r H2-H20-limit-plot-BA1D_2019, fig.height=5, fig.width=4}
plot_BA1D_2019_Eh <- BA1D_2019_profile_longer_Eh_T_P %>% ggplot(aes(x = depth_sampled_mbgl))+
  geom_line(aes(y = Eh_mV, color = parameter))+
  scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02))+
  scale_y_continuous(name = "Eh / [mV]")+
  coord_flip()+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "vertical")

plot_BA1D_2019_Eh
```

# Prepare data for facet plot

Join well logs

```{r join-well-logs}
# join by type column
# also taking this opportunity to add some metadata for plotting, and also remove data above the casing depths, which may be spurious and obscure true trends in the boreholes
# BA1_profiles <- full_join((BA1A_2018_profile %>%mutate(sampling_site = "BA1A", year_sampled = 2018, group_id = "BA1A_2018_log") %>% filter(depth_sampled_mbgl > 22)), (BA1A_2017_profile %>% mutate(sampling_site = "BA1A", year_sampled = 2017, group_id = "BA1A_2017_log") %>% filter(depth_sampled_mbgl > 22))) %>% full_join((BA1D_2018_profile  %>% mutate(sampling_site = "BA1D", year_sampled = 2018, group_id = "BA1D_2018_log") %>% filter(depth_sampled_mbgl > 26))) %>% full_join((BA1D_2019_profile  %>% mutate(sampling_site = "BA1D", year_sampled = 2019, group_id = "BA1D_2019_log") %>% filter(depth_sampled_mbgl > 26))) %>% arrange(depth_sampled_mbgl)

BA1_profiles <- (BA1A_2018_profile %>% mutate(sampling_site = "BA1A", year_sampled = 2018, group_id = "BA1A_2018_log") %>% filter(depth_sampled_mbgl > 22)) %>% full_join((BA1D_2019_profile  %>% mutate(sampling_site = "BA1D", year_sampled = 2019, group_id = "BA1D_2019_log") %>% filter(depth_sampled_mbgl > 26))) %>% arrange(depth_sampled_mbgl)

# print first few lines
head(BA1_profiles)
```

## Pivot longer for plotting on same scale

```{r pivot-longer-BA1}
BA1_profiles_longer <- BA1_profiles %>%
  select(sampling_site, year_sampled, depth_sampled_mbgl, pH, Eh_mV, group_id) %>%
  pivot_longer(cols = c(-sampling_site, -year_sampled, -depth_sampled_mbgl, -group_id), names_to = "parameter", values_to = "value", values_drop_na = TRUE)

BA1_profiles_longer %>% head()
```

## Add packer sampling data
```{r BA1-get-geochem}
# filter for data from only wells BA1A or BA1D
Oman_Geochem_BA1 <- Oman_Geochem %>% 
 filter(sampling_site == "BA1A" | sampling_site == "BA1D") %>% 
  # filter for 2019 only
  filter(year_sampled == 2019)

# print
Oman_Geochem_BA1
```
```{r BA1-pivot-longer}
# Get the relevant data and pivot it longer
Oman_Geochem_BA1_select <- Oman_Geochem_BA1 %>%
  # get only relevant data
  select(sampling_site, year_sampled, upper_packer_depth_mbct, lower_packer_depth_mbct, well_depth_mbgl, casing_height_magl, pH, Eh_mV)

# add a group ID for plotting purposes
Oman_Geochem_BA1_select <- Oman_Geochem_BA1_select %>% mutate(group_id = paste0(sampling_site, "_", year_sampled, "_", if_else(is.na(upper_packer_depth_mbct), 0, upper_packer_depth_mbct), "-", if_else(is.na(lower_packer_depth_mbct), well_depth_mbgl, lower_packer_depth_mbct)))

  # pivot it longer
Oman_Geochem_BA1_longer  <-  Oman_Geochem_BA1_select %>%
  pivot_longer(cols = c(-sampling_site, -year_sampled, -upper_packer_depth_mbct, -lower_packer_depth_mbct, -well_depth_mbgl, -casing_height_magl, -group_id), names_to = "parameter", values_to = "value", values_drop_na = TRUE)

# convert/calculate sample depths
Oman_Geochem_BA1_longer <- Oman_Geochem_BA1_longer %>% mutate(upper_packer_depth_mbgl = if_else(!is.na(upper_packer_depth_mbct), upper_packer_depth_mbct + casing_height_magl, 0), lower_packer_depth_mbgl = if_else(!is.na(lower_packer_depth_mbct), lower_packer_depth_mbct + casing_height_magl, well_depth_mbgl))

# pivot longer depths for plotting as lines graphs
Oman_Geochem_BA1_longer_depth <- Oman_Geochem_BA1_longer %>%
  select(-upper_packer_depth_mbct, -lower_packer_depth_mbct, -well_depth_mbgl, -casing_height_magl) %>%
  pivot_longer(cols = c(-sampling_site, -year_sampled, -parameter, -value, -group_id), names_to = "depth_type", values_to = "depth_sampled_mbgl", values_drop_na = TRUE) %>%
  select(sampling_site, year_sampled, group_id, depth_sampled_mbgl, parameter, value)

Oman_Geochem_BA1_longer_depth
```
```{r BA1-add-sampling-method}
BA1_profiles_and_packer_geochem <- full_join((BA1_profiles_longer %>% mutate(sample_method = "Wireline logging")), (Oman_Geochem_BA1_longer_depth %>% mutate(sample_method = "Packer")))
```


Prepping data to plot and label correctly
```{r prepare-BA1-plot-labels}
# set factors to set order of parameters on plot
BA1_profiles_and_packer_geochem$parameter <- factor(BA1_profiles_and_packer_geochem$parameter, levels = c("pH", "Eh_mV"))

# rename how parameters should be displayed on plot using LaTeX
BA1_profiles_and_packer_geochem <- BA1_profiles_and_packer_geochem %>%
  mutate(
    tex_params = as_factor(parameter) %>%
      fct_recode(
        # "T / $\\lbrack$$\\degree$C$\\rbrack$" = "T_C",
        "pH" = "pH",
        "$\\textit{Eh}$ / $\\lbrack$mV$\\rbrack$" = "Eh_mV"
  )
  )
```

## Plot
```{r BA1_profiles_plot, fig.width=8, fig.height=5}
BA1_profiles_plot <- BA1_profiles_and_packer_geochem %>% ggplot(aes(
  x = depth_sampled_mbgl,
  # linetype = as.character(year_sampled),
  color = sampling_site,
  # linetype = sample_method,
  size = sample_method,
  group = group_id
  )) +
  
  # add line graph for applicable parameters
  geom_line(data = (BA1_profiles_and_packer_geochem %>% filter(parameter == "pH" | parameter == "Eh_mV")), aes(y = value)) +

  # flip coordinates (this makes it possible to do vertical line graph)
    coord_flip() +
  
  # plot styling
  scale_color_manual(name = "Well", values = c("BA1A" = "#E69F00", "BA1D" = "#0072B2"))+
  scale_size_manual(name = "Method", values = c("Wireline logging" = 0.3, "Packer" = 3)) +
  scale_x_reverse(name = "Depth / [m]", limits = c(410, -2), expand = c(0, 0)) + 
  scale_y_continuous(name = NULL, sec.axis = dup_axis()) +
  
  # generate plot facets
  facet_rep_grid(cols = vars(tex_params), scales = "free", labeller = latex_labeller)+
  
    # plot design
  theme_classic(base_size = 18) +
  theme(
    legend.position = "none",
    # legend.direction = "vertical",
    panel.grid = element_blank(),
    plot.margin = margin(1,1,1,1),
    strip.placement = "outside",
    axis.text.x.bottom = element_blank(),
    axis.ticks.x.bottom = element_blank(),
    axis.line.x.bottom = element_blank(),
    axis.line.y.left = element_line(),
    strip.background = element_blank(),
    strip.text.y = element_blank(),
    panel.spacing = unit(-5, "pt")
  )


BA1_profiles_plot
```
