---
title: "BA1 geophysical/chemical logs"
subtitle: "Source file: 201012_BA1_LOGS.Rmd"
author: "Daniel Nothaft"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged # omit to disable paged data table output
    css: stylesheet.css # omit if no need for custom stylesheet
    number_sections: yes # change to no for unnumbered sections
    toc: yes # change to no to disable table of contents
    toc_float: true # change to false to keep toc at the top
    toc_depth: 3 # change to specify which headings to include in toc
    code_folding: show # change to hide to hide code by default
editor_options:
  chunk_output_type: inline
---
# Setup
## Load code libraries
```{r setup, message=FALSE, warning=FALSE}
# load libraries
library(tidyverse) # dplyr, tidyr, ggplot
library(readxl) # reading excel files
library(knitr) # generating reports
library(ggpmisc) # add equations of best fit to ggplot
library(scales) # making log scales with exponents
library(lemon) # repeat axis lines for paneled figures in ggplot

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"), fig.keep = "all",
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.path = file.path("fig_output/", paste0(gsub("\\.[Rr]md", "/", knitr::current_input()))),
  cache.path = file.path("cache/", paste0(gsub("\\.[Rr]md", "/", knitr::current_input())))
)
```

```{r source}
# source all relevant scripting files
source(file.path("scripts", "plotting_functions.R"))
```

## Load data

```{r load-data}
# read BA1A borehole log from March 16th, 2018
BA1A_2018_profile_raw <- read_xlsx("data_raw/Summary_metadata/BA1A_QL40IDRO_20180316_DOWN_5cm.xlsx", na = c("-99999.00", "-99999"))

# read BA1A borehole log from April 22nd, 2017
BA1A_2017_profile_raw <- read_xlsx("data_raw/Summary_metadata/BA1A_Muset_22APRIL2017_run2_pass1_dw_5cm.xlsx", na = c("-99999.00", "-99999"))

# read BA1A borehole log from March 16th, 2018
BA1D_2018_profile_raw <- read_xlsx("data_raw/Summary_metadata/BA1D_QL40IDRO_20180319_down_5cm.xlsx", na = c("-99999.00", "-99999"))

# Read in conversion from Ag/AgCl electrode to SHE as function of T
Ag_AgCl_E_T <- read_xlsx("data_raw/Summary_metadata/Ag_AgCl_E_T.xlsx", na = "NA")
```

# BA1A 2018 well log

## clean up data
```{r get-relevant-log-data-BA1A-2018}
# select relevant columns and rename them
BA1A_2018_profile <- BA1A_2018_profile_raw %>% select(depth_sampled_mbgl = depth_m, P_dbar = pressure, T_C = Temp_C, pH, ORP_mV_Ag_AgCl = ORP, cond_uS_per_cm = Cond_FW)

# Filter out nonsense values (i.e. when probe is above water table)
BA1A_2018_profile <- BA1A_2018_profile %>% filter_all(all_vars(!is.na(.)))

# convert conductivity to mS/cm
BA1A_2018_profile <- BA1A_2018_profile %>% mutate(cond_mS_per_cm = cond_uS_per_cm * 1e-3) %>% select(-cond_uS_per_cm)

# print some summary statistics
BA1A_2018_profile %>% summary()

BA1A_2018_profile %>% head()
```

```{r subset-and-print-log-BA1A_2018}
# subset the data for readability and print
BA1A_2018_profile %>% slice(seq(from = 1, to = 10000, by = 200)) %>% kable(digits=1)
```

## Convert to ORP to SHE

Plot conversion between Ag/AgCl electrode and SHE as function of temperature
```{r plot-Ag-AgCl-SHE-conversion}
formula <- Ag_AgCl_E_T$E_Ag_AgCl_KCl_sat_mV ~ Ag_AgCl_E_T$T_C

plot_redox_offset_T <- Ag_AgCl_E_T %>% ggplot(aes(x = T_C, y = E_Ag_AgCl_KCl_sat_mV))+
  geom_point()+
  geom_smooth(method="lm")+
  stat_poly_eq(aes(label =  paste(stat(eq.label), stat(rr.label), sep = "~~~~")),
               formula = formula, parse = TRUE, rr.digits = 2, label.x = .2, label.y = .9)+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank())

plot_redox_offset_T
```
Save a linear model of the relationship between Ag/AgCl electrode and SHE as function of temperature
```{r save-lm-Ag-AgCl-SHE}
Ag_AgCl_E_T_lm <- lm(formula)

Ag_AgCl_E_T_lm_summ <- Ag_AgCl_E_T_lm %>% summary()

Ag_AgCl_E_T_lm_summ
```
Save slope
```{r save-lm-Ag-AgCl-SHE-slope}
Ag_AgCl_E_T_lm_slope <- coef(Ag_AgCl_E_T_lm_summ)["Ag_AgCl_E_T$T_C",  "Estimate"]

Ag_AgCl_E_T_lm_slope
```
Save intercept
```{r save-lm-Ag-AgCl-SHE-int}
Ag_AgCl_E_T_lm_int <- coef(Ag_AgCl_E_T_lm_summ)["(Intercept)",  "Estimate"]

Ag_AgCl_E_T_lm_int
```

Apply conversion
```{r apply-Ag-AgCl-SHE-conversion-NSHQ14}
BA1A_2018_profile <- BA1A_2018_profile %>% mutate(Eh_mV = ORP_mV_Ag_AgCl + Ag_AgCl_E_T_lm_int + Ag_AgCl_E_T_lm_slope * T_C) # convert probe ORP_mV_Ag_AgCl reading from Ag/AgCl to SHE

# print some summary statistics
BA1A_2018_profile %>% select(Eh_mV) %>% summary()
```

## Hydrostatic pressure
Pressure will be used to assess calculated stability limit of water with respect to molecular hydrogen.

Calculate from theory
```{r calc-hydrostatic-P-NSHQ14}
P_BA1A_2018_atmospheric_Pa <- 95490 # measured ambient pressure with Garmin GPSMAP 64S by DBN in 2019
P_BA1A_2018_atmospheric_bar <-  P_BA1A_2018_atmospheric_Pa * 1e-5

water_level_BA1A_2018_m <- as.numeric(BA1A_2018_profile %>% select(depth_sampled_mbgl) %>% min())

BA1A_2018_profile <- BA1A_2018_profile %>% mutate(water_column_above_m = depth_sampled_mbgl - water_level_BA1A_2018_m)


# Hydrostatic pressure in a liquid can be calculated as
# 
# p = ρ g h                         (1)
# 
# where
# 
# p = pressure in liquid (N/m2, Pa, lbf/ft2, psf)
# 
# ρ = density of liquid (kg/m3, slugs/ft3)
# 
# g = acceleration of gravity (9.81 m/s2, 32.17405 ft/s2)
# 
# h = height of fluid column - or depth in the fluid where pressure is measured (m, ft)

# https://www.engineeringtoolbox.com/hydrostatic-pressure-water-d_1632.html

density_water <- 1000 # kg/m3
g = 9.81 # acceleration of gravity, m/s2


BA1A_2018_profile <- BA1A_2018_profile %>% mutate(P_hydrostatic_calc_Pa = P_BA1A_2018_atmospheric_Pa + density_water * g * water_column_above_m)

BA1A_2018_profile <- BA1A_2018_profile %>% mutate(P_hydrostatic_calc_bar = P_hydrostatic_calc_Pa * 1e-5)

# print some summary statistics
BA1A_2018_profile %>% select(P_hydrostatic_calc_bar) %>% summary()
```

```{r convert-hydrostatic-P-to-bar-BA1A_2018}
# convert measured pressure to bar
BA1A_2018_profile <- BA1A_2018_profile %>% mutate(P_bar = P_dbar / 10)
```

```{r prepare-P-plot-BA1A_2018}
# prepare for plotting
BA1A_2018_profile_longer_P <- BA1A_2018_profile %>% select(depth_sampled_mbgl, P_hydrostatic_calc_bar, P_bar) %>% gather(-depth_sampled_mbgl, key = parameter, value = P_bar) %>% mutate(type = if_else(parameter == "P_hydrostatic_calc_bar", "calculated", "measured"))
```

Compare theoretical hydrostatic pressure versus measured pressure
```{r P-plot-BA1A_2018, fig.height=4, fig.width=4}
plot_BA1A_2018_P <- BA1A_2018_profile_longer_P %>% ggplot(aes(x = depth_sampled_mbgl))+
  geom_line(aes(y = P_bar, color = type))+
  scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02), breaks = c(0, 50, 100, 150, 200, 250, 300))+
    scale_y_continuous(name = latex2exp::TeX("$\\textit{P}_{hydrostatic}\\,/\\,\\lbrack bar \\rbrack$"))+
  coord_flip()+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank())

plot_BA1A_2018_P
```
They are similar.

## Stability limits of water

Eh, f(T,P)
```{r calc-H2-H20-limit-BA1A_2018}
Faraday <- 96485.33212331 # C/mol
R <- 8.31446261815324 # J/(mol⋅K) gas constant

BA1A_2018_profile <- BA1A_2018_profile %>% mutate(T_K= T_C + 273.15) # convert T to Kelvins

# calculate H20-H2 stability limit
BA1A_2018_profile <- BA1A_2018_profile %>% mutate(H2O_H2_limit_Eh_f_T_P_mV = (-R*T_K/2/Faraday * log(((10^-pH)^-2)*P_bar))*1000)

BA1A_2018_profile %>% select(H2O_H2_limit_Eh_f_T_P_mV) %>% summary()
```

```{r prepare-H2-H20-limit-plot-BA1A_2018}
# prepare for plotting
BA1A_2018_profile_longer_Eh_T_P <- BA1A_2018_profile %>% select(depth_sampled_mbgl, Eh_mV, H2O_H2_limit_Eh_f_T_P_mV) %>% gather(-depth_sampled_mbgl, key = parameter, value = Eh_mV)
```

```{r H2-H20-limit-plot-BA1A_2018, fig.height=5, fig.width=4}
plot_BA1A_2018_Eh <- BA1A_2018_profile_longer_Eh_T_P %>% ggplot(aes(x = depth_sampled_mbgl))+
  geom_line(aes(y = Eh_mV, color = parameter))+
  scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02))+
  scale_y_continuous(name = "Eh / [mV]")+
  coord_flip()+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "vertical")

plot_BA1A_2018_Eh
```

# BA1A 2017 well log

## clean up data
```{r get-relevant-log-data-BA1A_2017}
# select relevant columns and rename them
BA1A_2017_profile <- BA1A_2017_profile_raw %>% select(depth_sampled_mbgl = `DEPT[M]`, P_dbar = PRESSURE, T_C = TEMP, pH = PH, ORP_mV_Ag_AgCl = EH, cond_mS_per_cm = COND)

# Filter out nonsense values (i.e. when probe is above water table)
BA1A_2017_profile <- BA1A_2017_profile %>% filter(!is.na(ORP_mV_Ag_AgCl))

# print some summary statistics
BA1A_2017_profile %>% summary()
```

```{r subset-and-print-log-BA1A_2017}
# subset the data for readability and print
BA1A_2017_profile %>% slice(seq(from = 1, to = 40000, by = 1000)) %>% kable(digits=1)
```

## Convert to ORP to SHE

Apply conversion
```{r apply-Ag-AgCl-SHE-conversion-BA1A_2017}
BA1A_2017_profile <- BA1A_2017_profile %>% mutate(Eh_mV = ORP_mV_Ag_AgCl + Ag_AgCl_E_T_lm_int + Ag_AgCl_E_T_lm_slope * T_C) # convert probe ORP_mV_Ag_AgCl reading from Ag/AgCl to SHE

# print some summary statistics
BA1A_2017_profile %>% select(Eh_mV) %>% summary()
```

## Hydrostatic pressure
Pressure will be used to assess calculated stability limit of water with respect to molecular hydrogen.

Calculate from theory
```{r calc-hydrostatic-P-BA1A_2017}
P_BA1A_2017_atmospheric_Pa <- 95490 # measured ambient pressure with Garmin GPSMAP 64S by DBN in 2019
P_BA1A_2017_atmospheric_bar <-  P_BA1A_2017_atmospheric_Pa * 1e-5

water_level_BA1A_2017_m <- as.numeric(BA1A_2017_profile %>% select(depth_sampled_mbgl) %>% min())

BA1A_2017_profile <- BA1A_2017_profile %>% mutate(water_column_above_m = depth_sampled_mbgl - water_level_BA1A_2017_m)


# Hydrostatic pressure in a liquid can be calculated as
# 
# p = ρ g h                         (1)
# 
# where
# 
# p = pressure in liquid (N/m2, Pa, lbf/ft2, psf)
# 
# ρ = density of liquid (kg/m3, slugs/ft3)
# 
# g = acceleration of gravity (9.81 m/s2, 32.17405 ft/s2)
# 
# h = height of fluid column - or depth in the fluid where pressure is measured (m, ft)

# https://www.engineeringtoolbox.com/hydrostatic-pressure-water-d_1632.html

density_water <- 1000 # kg/m3
g = 9.81 # acceleration of gravity, m/s2


BA1A_2017_profile <- BA1A_2017_profile %>% mutate(P_hydrostatic_calc_Pa = P_BA1A_2017_atmospheric_Pa + density_water * g * water_column_above_m)

BA1A_2017_profile <- BA1A_2017_profile %>% mutate(P_hydrostatic_calc_bar = P_hydrostatic_calc_Pa * 1e-5)

# print some summary statistics
BA1A_2017_profile %>% select(P_hydrostatic_calc_bar) %>% summary()
```
```{r convert-hydrostatic-P-to-bar-BA1A_2017}
# convert measured pressure to bar
BA1A_2017_profile <- BA1A_2017_profile %>% mutate(P_bar = P_dbar / 10)
```

```{r prepare-P-plot-BA1A_2017}
# prepare for plotting
BA1A_2017_profile_longer_P <- BA1A_2017_profile %>% select(depth_sampled_mbgl, P_hydrostatic_calc_bar, P_bar) %>% gather(-depth_sampled_mbgl, key = parameter, value = P_bar) %>% mutate(type = if_else(parameter == "P_hydrostatic_calc_bar", "calculated", "measured"))
```

Compare theoretical hydrostatic pressure versus measured pressure
```{r P-plot-BA1A_2017, fig.height=4, fig.width=4}
plot_BA1A_2017_P <- BA1A_2017_profile_longer_P %>% ggplot(aes(x = depth_sampled_mbgl))+
  geom_line(aes(y = P_bar, color = type))+
  scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02), breaks = c(0, 50, 100, 150, 200, 250, 300))+
    scale_y_continuous(name = latex2exp::TeX("$\\textit{P}_{hydrostatic}\\,/\\,\\lbrack bar \\rbrack$"))+
  coord_flip()+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank())

plot_BA1A_2017_P
```
They are similar.

## Stability limits of water

Eh, f(T,P)
```{r calc-H2-H20-limit-BA1A_2017}
Faraday <- 96485.33212331 # C/mol
R <- 8.31446261815324 # J/(mol⋅K) gas constant

BA1A_2017_profile <- BA1A_2017_profile %>% mutate(T_K= T_C + 273.15) # convert T to Kelvins

# calculate H20-H2 stability limit
BA1A_2017_profile <- BA1A_2017_profile %>% mutate(H2O_H2_limit_Eh_f_T_P_mV = (-R*T_K/2/Faraday * log(((10^-pH)^-2)*P_bar))*1000)

BA1A_2017_profile %>% select(H2O_H2_limit_Eh_f_T_P_mV) %>% summary()
```

```{r prepare-H2-H20-limit-plot-BA1A_2017}
# prepare for plotting
BA1A_2017_profile_longer_Eh_T_P <- BA1A_2017_profile %>% select(depth_sampled_mbgl, Eh_mV, H2O_H2_limit_Eh_f_T_P_mV) %>% gather(-depth_sampled_mbgl, key = parameter, value = Eh_mV)
```

```{r H2-H20-limit-plot-BA1A_2017, fig.height=5, fig.width=4}
plot_BA1A_2017_Eh <- BA1A_2017_profile_longer_Eh_T_P %>% ggplot(aes(x = depth_sampled_mbgl))+
  geom_line(aes(y = Eh_mV, color = parameter))+
  scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02))+
  scale_y_continuous(name = "Eh / [mV]")+
  coord_flip()+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "vertical")

plot_BA1A_2017_Eh
```

# BA1D 2018 well log

## clean up data
```{r get-relevant-log-data-BA1D_2018}
# select relevant columns and rename them
BA1D_2018_profile <- BA1D_2018_profile_raw %>% select(depth_sampled_mbgl = `DEPT[M]`, P_dbar = PRESSURE, T_C = TEMP, pH = PH, ORP_mV_Ag_AgCl = EH, cond_mS_per_cm = COND)

# Filter out nonsense values (i.e. when probe is above water table)
BA1D_2018_profile <- BA1D_2018_profile %>% filter(!is.na(ORP_mV_Ag_AgCl))

# print some summary statistics
BA1D_2018_profile %>% summary()
```

```{r subset-and-print-log-BA1D_2018}
# subset the data for readability and print
BA1D_2018_profile %>% slice(seq(from = 1, to = 40000, by = 1000)) %>% kable(digits=1)
```

## Convert to ORP to SHE

Apply conversion
```{r apply-Ag-AgCl-SHE-conversion-BA1D_2018}
BA1D_2018_profile <- BA1D_2018_profile %>% mutate(Eh_mV = ORP_mV_Ag_AgCl + Ag_AgCl_E_T_lm_int + Ag_AgCl_E_T_lm_slope * T_C) # convert probe ORP_mV_Ag_AgCl reading from Ag/AgCl to SHE

# print some summary statistics
BA1D_2018_profile %>% select(Eh_mV) %>% summary()
```

## Hydrostatic pressure
Pressure will be used to assess calculated stability limit of water with respect to molecular hydrogen.

Calculate from theory
```{r calc-hydrostatic-P-BA1D_2018}
P_BA1D_2018_atmospheric_Pa <- 95490 # measured ambient pressure with Garmin GPSMAP 64S by DBN in 2019
P_BA1D_2018_atmospheric_bar <-  P_BA1D_2018_atmospheric_Pa * 1e-5

water_level_BA1D_2018_m <- as.numeric(BA1D_2018_profile %>% select(depth_sampled_mbgl) %>% min())

BA1D_2018_profile <- BA1D_2018_profile %>% mutate(water_column_above_m = depth_sampled_mbgl - water_level_BA1D_2018_m)


# Hydrostatic pressure in a liquid can be calculated as
# 
# p = ρ g h                         (1)
# 
# where
# 
# p = pressure in liquid (N/m2, Pa, lbf/ft2, psf)
# 
# ρ = density of liquid (kg/m3, slugs/ft3)
# 
# g = acceleration of gravity (9.81 m/s2, 32.17405 ft/s2)
# 
# h = height of fluid column - or depth in the fluid where pressure is measured (m, ft)

# https://www.engineeringtoolbox.com/hydrostatic-pressure-water-d_1632.html

density_water <- 1000 # kg/m3
g = 9.81 # acceleration of gravity, m/s2


BA1D_2018_profile <- BA1D_2018_profile %>% mutate(P_hydrostatic_calc_Pa = P_BA1D_2018_atmospheric_Pa + density_water * g * water_column_above_m)

BA1D_2018_profile <- BA1D_2018_profile %>% mutate(P_hydrostatic_calc_bar = P_hydrostatic_calc_Pa * 1e-5)

# print some summary statistics
BA1D_2018_profile %>% select(P_hydrostatic_calc_bar) %>% summary()
```
```{r convert-hydrostatic-P-to-bar-BA1D_2018}
# convert measured pressure to bar
BA1D_2018_profile <- BA1D_2018_profile %>% mutate(P_bar = P_dbar / 10)
```

```{r prepare-P-plot-BA1D_2018}
# prepare for plotting
BA1D_2018_profile_longer_P <- BA1D_2018_profile %>% select(depth_sampled_mbgl, P_hydrostatic_calc_bar, P_bar) %>% gather(-depth_sampled_mbgl, key = parameter, value = P_bar) %>% mutate(type = if_else(parameter == "P_hydrostatic_calc_bar", "calculated", "measured"))
```

Compare theoretical hydrostatic pressure versus measured pressure
```{r P-plot-BA1D_2018, fig.height=4, fig.width=4}
plot_BA1D_2018_P <- BA1D_2018_profile_longer_P %>% ggplot(aes(x = depth_sampled_mbgl))+
  geom_line(aes(y = P_bar, color = type))+
  scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02), breaks = c(0, 50, 100, 150, 200, 250, 300))+
    scale_y_continuous(name = latex2exp::TeX("$\\textit{P}_{hydrostatic}\\,/\\,\\lbrack bar \\rbrack$"))+
  coord_flip()+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank())

plot_BA1D_2018_P
```
They are similar.

## Stability limits of water

Eh, f(T,P)
```{r calc-H2-H20-limit-BA1D_2018}
Faraday <- 96485.33212331 # C/mol
R <- 8.31446261815324 # J/(mol⋅K) gas constant

BA1D_2018_profile <- BA1D_2018_profile %>% mutate(T_K= T_C + 273.15) # convert T to Kelvins

# calculate H20-H2 stability limit
BA1D_2018_profile <- BA1D_2018_profile %>% mutate(H2O_H2_limit_Eh_f_T_P_mV = (-R*T_K/2/Faraday * log(((10^-pH)^-2)*P_bar))*1000)

BA1D_2018_profile %>% select(H2O_H2_limit_Eh_f_T_P_mV) %>% summary()
```

```{r prepare-H2-H20-limit-plot-BA1D_2018}
# prepare for plotting
BA1D_2018_profile_longer_Eh_T_P <- BA1D_2018_profile %>% select(depth_sampled_mbgl, Eh_mV, H2O_H2_limit_Eh_f_T_P_mV) %>% gather(-depth_sampled_mbgl, key = parameter, value = Eh_mV)
```

```{r H2-H20-limit-plot-BA1D_2018, fig.height=5, fig.width=4}
plot_BA1D_2018_Eh <- BA1D_2018_profile_longer_Eh_T_P %>% ggplot(aes(x = depth_sampled_mbgl))+
  geom_line(aes(y = Eh_mV, color = parameter))+
  scale_x_reverse(name = "Depth / [mbgl]", expand = c(0.03,0.02))+
  scale_y_continuous(name = "Eh / [mV]")+
  coord_flip()+
  theme_bw(base_size = 12)+
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "vertical")

plot_BA1D_2018_Eh
```

# Prepare data for facet plot
Get relevant carbonate mineral isotope data
```{r get-carb-min-iso-data}
data_NSHQ14_BA3A <- data_carb %>% select(sampling_site, Fm, depth_sampled_mbgl, carbonate_d13C_VPDB_permil)

# print
data_NSHQ14_BA3A
```

convert conductivity units
```{r convert-cond-units}
NSHQ14_profile <- NSHQ14_profile %>% mutate(cond_mS_cm = cond_uS_per_cm/1000) %>% select(-cond_uS_per_cm) %>% mutate(sampling_site = "NSHQ14")
BA3A_profile <- BA3A_profile %>% mutate(cond_mS_cm = cond_uS_per_cm/1000) %>% select(-cond_uS_per_cm) %>% mutate(sampling_site = "BA3A")
```

Join well logs, and carbonate vein and isotope data

```{r join-well-logs-iso-and-vein-data}
# join by type column
NSHQ14_BA3A_profiles_and_isotopes <- full_join(NSHQ14_profile, data_NSHQ14_BA3A) %>% full_join(BA3A_profile) %>% full_join(BA3A_vein_types_carb) %>% arrange(depth_sampled_mbgl)

# print first few lines
head(NSHQ14_BA3A_profiles_and_isotopes)
```

## Pivot longer for plotting on same scale

```{r pivot-longer-NSH14-BA3A}
NSHQ14_BA3A_profiles_and_isotopes_longer <- NSHQ14_BA3A_profiles_and_isotopes %>% select(-P_dbar, -ORP_mV_Ag_AgCl, -water_column_above_m, -P_hydrostatic_calc_Pa, -P_hydrostatic_calc_bar, -P_bar, -T_K, -H2O_H2_limit_Eh_f_T_P_mV) %>% pivot_longer(cols = c(-sampling_site, -depth_sampled_mbgl), names_to = "parameter", values_to = "value", values_drop_na = TRUE)

NSHQ14_BA3A_profiles_and_isotopes_longer
```

Prepping data to plot and label correctly
```{r prepare-NSHQ14-BA3A-plot-labels}
# set factors to set order of parameters on plot
NSHQ14_BA3A_profiles_and_isotopes_longer$parameter <- factor(NSHQ14_BA3A_profiles_and_isotopes_longer$parameter, levels = c("T_C", "pH", "Eh_mV", "cond_mS_cm", "vein_type_carbonate", "carbonate_d13C_VPDB_permil", "Fm"))

# rename how parameters should be displayed on plot using LaTeX
NSHQ14_BA3A_profiles_and_isotopes_longer <- NSHQ14_BA3A_profiles_and_isotopes_longer %>%
  mutate(
    tex_params_overset = as_factor(parameter)%>%
      fct_recode(
        "T / $\\lbrack$$\\degree$C$\\rbrack$" = "T_C",
        "pH" = "pH",
        "Eh / $\\lbrack$mV$\\rbrack$" = "Eh_mV",
        "$\\overset{Conductivity}{/\\,$\\lbrack$mS$\\cdot$cm^{-1}$\\rbrack$}$" = "cond_mS_cm",
        "$\\overset{Veins\\,with}{carbonate,\\,BA3A}$" = "vein_type_carbonate",
        "$\\overset{Carbonate}{\\delta^{13}C\\,/\\,$\\lbrack$‰\\,VPDB\\rbrack}$" = "carbonate_d13C_VPDB_permil",
        "$\\overset{Carbonate}{fraction\\,modern}$" = "Fm"
  )
  )
```

## Plot
```{r NSHQ14-BA3A-downhole-plot-tex-overset, fig.width=7, fig.height=4}
NSHQ14_BA3A_profiles_and_isotopes_plot <- NSHQ14_BA3A_profiles_and_isotopes_longer %>% ggplot(aes(
  x = depth_sampled_mbgl,
  fill = sampling_site,
  color = sampling_site
  )) +
  
  # add line graph for applicable parameters
  geom_line(data = (NSHQ14_BA3A_profiles_and_isotopes_longer %>% filter(parameter == "T_C" | parameter == "pH" | parameter == "Eh_mV" | parameter == "cond_mS_cm")), aes(y = value)) +
  
  # add scatterplot for applicable parameters
  geom_point(data = (NSHQ14_BA3A_profiles_and_isotopes_longer %>% filter(parameter == "Fm" | parameter == "carbonate_d13C_VPDB_permil")), aes(y = value)) + 

  # add histogram for applicable parameter
  geom_histogram(data = (NSHQ14_BA3A_profiles_and_isotopes_longer %>% filter(parameter == "vein_type_carbonate")), binwidth = 5, boundary = 0, color = NA) +

  # flip coordinates (this makes it possible to do vertical line graph)
    coord_flip() +
  
  # plot styling
  scale_fill_manual(name = "Well", values = c("NSHQ14" = "black", "BA3A" = "#0072B2"))+
  scale_color_manual(name = "Well", values = c("NSHQ14" = "black", "BA3A" = "#0072B2"))+
  scale_x_reverse(name = "Depth / [m]", limits = c(300, -2), expand = c(0, 0), breaks = c(0, 50, 100, 150, 200, 250, 300)) + 
  scale_y_continuous(name = NULL) +
  
  # generate plot facets
  # facet_grid(cols = vars(params_recode), scales = "free", switch = "y")+
  # for alternative names
    facet_grid(cols = vars(tex_params_overset), scales = "free", labeller = latex_labeller, switch = "y")+

  # plot design
  theme_bw(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = -90, vjust = 0.5, hjust = 0),
    legend.position = "bottom",
    legend.direction = "horizontal",
    panel.grid = element_blank(),
    plot.margin = margin(1,1,0,1)
  )

NSHQ14_BA3A_profiles_and_isotopes_plot
```

```{r}
equal_breaks <- function(n = 3, s = 0.05, ...){
  function(x){
    # rescaling
    d <- s * diff(range(x)) / (1+2*s)
    seq(min(x)+d, max(x)-d, length=n)
  }
}
```


```{r NSHQ14-BA3A-downhole-plot-tex-overset-clean, fig.width=7, fig.height=4}
NSHQ14_BA3A_profiles_and_isotopes_plot_clean <- NSHQ14_BA3A_profiles_and_isotopes_longer %>% ggplot(aes(
  x = depth_sampled_mbgl,
  fill = sampling_site,
  color = sampling_site
  )) +
  
  # add line graph for applicable parameters
  geom_line(data = (NSHQ14_BA3A_profiles_and_isotopes_longer %>% filter(parameter == "T_C" | parameter == "pH" | parameter == "Eh_mV" | parameter == "cond_mS_cm")), aes(y = value)) +
  
  # add scatterplot for applicable parameters
  geom_point(data = (NSHQ14_BA3A_profiles_and_isotopes_longer %>% filter(parameter == "Fm" | parameter == "carbonate_d13C_VPDB_permil")), aes(y = value)) + 

  # add histogram for applicable parameter
  geom_histogram(data = (NSHQ14_BA3A_profiles_and_isotopes_longer %>% filter(parameter == "vein_type_carbonate")), binwidth = 5, boundary = 0, color = NA) +

  # flip coordinates (this makes it possible to do vertical line graph)
    coord_flip() +
  
  # plot styling
  scale_fill_manual(name = "Well", values = c("NSHQ14" = "#56B4E9", "BA3A" = "black"))+
  scale_color_manual(name = "Well", values = c("NSHQ14" = "#56B4E9", "BA3A" = "black"))+
  scale_x_reverse(name = "Depth / [m]", limits = c(300, -2), expand = c(0, 0), breaks = c(0, 50, 100, 150, 200, 250, 300)) + 
  scale_y_continuous(name = NULL, sec.axis = dup_axis(), breaks = breaks_extended(n=3.5)) +
  
  # generate plot facets
  # facet_grid(cols = vars(params_recode), scales = "free", switch = "y")+
  # for alternative names
    # facet_grid(cols = vars(tex_params_overset), scales = "free", labeller = latex_labeller, switch = "y")+
  facet_rep_grid(cols = vars(tex_params_overset), scales = "free", labeller = latex_labeller, switch = "y")+
  
  # plot design
  theme_classic(base_size = 8.5) +
  theme(
    axis.text.x.top = element_text(size = 6.5),
    axis.text.y = element_text(size = 6.5),
    legend.position = "bottom",
    legend.direction = "horizontal",
    panel.grid = element_blank(),
    plot.margin = margin(0,6,0,1),
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text = element_text(vjust = 0, size = 8),
    axis.text.x.bottom = element_blank(),
    axis.ticks.x.bottom = element_blank(),
    axis.line.x.bottom = element_blank(),
    axis.line.y.left = element_line(),
    panel.spacing = unit(-8, "pt"),
    axis.title.y = element_text(size = 7.5)
  )

NSHQ14_BA3A_profiles_and_isotopes_plot_clean
```

```{r NSHQ14-BA3A-downhole-plot-tex-overset-no-side-tags}
NSHQ14_BA3A_profiles_and_isotopes_plot_clean %>% tag_facet_outside_keep_strip_no_side_tags(
  open = "",
  close = "",
  tag_fun_top = function(i) LETTERS[i]
)
```


# Geotherm

## NSHQ14

```{r geotherm-NSHQ14-full, warning=FALSE}
plot_T_NSHQ14 <- NSHQ14_profile %>% ggplot( mapping = aes(
    x = T_C,
    y = depth_sampled_mbgl)) +
  geom_point() +
    geom_smooth(method="lm", color = "blue") +
  stat_poly_eq(aes(label =  paste(stat(eq.label), stat(rr.label), sep = "~~~~")),
               formula = NSHQ14_profile$T_C ~ NSHQ14_profile$depth_sampled_mbgl, parse = TRUE, rr.digits = 2, color = "blue", label.x = 0.2)+
  scale_y_reverse(name = latex2exp::TeX("depth$\\,$/$\\,$$\\left[$m$\\right]$"))+
    scale_x_continuous(name = latex2exp::TeX("temperature$\\,$/$\\,$$\\left[$˚C$\\right]$"))+
  theme_bw()+
  theme(
    panel.grid = element_blank()
  )

plot_T_NSHQ14 
```

```{r geotherm-NSHQ14-below-100m, warning=FALSE}
NSHQ14_profile_below_100m <- NSHQ14_profile %>% filter(depth_sampled_mbgl > 100)

plot_T_NSHQ14_below_100m <- NSHQ14_profile_below_100m %>% ggplot( mapping = aes(
    x = T_C,
    y = depth_sampled_mbgl)) +
  geom_point() +
    geom_smooth(method="lm", color = "blue") +
  stat_poly_eq(aes(label =  paste(stat(eq.label), stat(rr.label), sep = "~~~~")),
               formula = NSHQ14_profile_below_100m$T_C ~ NSHQ14_profile_below_100m$depth_sampled_mbgl, parse = TRUE, rr.digits = 2, color = "blue", label.x = 0.2)+
  scale_y_reverse(name = latex2exp::TeX("depth$\\,$/$\\,$$\\left[$m$\\right]$"))+
    scale_x_continuous(name = latex2exp::TeX("temperature$\\,$/$\\,$$\\left[$˚C$\\right]$"))+
    theme_bw()+
  theme(
    panel.grid = element_blank()
  )

plot_T_NSHQ14_below_100m 
```

```{r lm-T-d-NSHQ14-below-100m}
# generate linear model
model_T_d_NSHQ14_below_100m = lm(T_C ~ depth_sampled_mbgl,
           data = NSHQ14_profile_below_100m)

# print model summary
summary(model_T_d_NSHQ14_below_100m)
```

## BA3A

```{r geotherm-BA3A-full, warning=FALSE}
plot_T_BA3A <- BA3A_profile %>% ggplot( mapping = aes(
    x = T_C,
    y = depth_sampled_mbgl)) +
  geom_point() +
    geom_smooth(method="lm", color = "blue") +
  stat_poly_eq(aes(label =  paste(stat(eq.label), stat(rr.label), sep = "~~~~")),
               formula = BA3A_profile$T_C ~ BA3A_profile$depth_sampled_mbgl, parse = TRUE, rr.digits = 2, color = "blue", label.x = 0.2)+
  scale_y_reverse(name = latex2exp::TeX("depth$\\,$/$\\,$$\\left[$m$\\right]$"))+
    scale_x_continuous(name = latex2exp::TeX("temperature$\\,$/$\\,$$\\left[$˚C$\\right]$"))+
  theme_bw()+
  theme(
    panel.grid = element_blank()
  )

plot_T_BA3A 
```

```{r geotherm-BA3A-below-100m, warning=FALSE}
BA3A_profile_below_100m <- BA3A_profile %>% filter(depth_sampled_mbgl > 100)

plot_T_BA3A_below_100m <- BA3A_profile_below_100m %>% ggplot( mapping = aes(
    x = T_C,
    y = depth_sampled_mbgl)) +
  geom_point() +
    geom_smooth(method="lm", color = "blue") +
  stat_poly_eq(aes(label =  paste(stat(eq.label), stat(rr.label), sep = "~~~~")),
               formula = BA3A_profile_below_100m$T_C ~ BA3A_profile_below_100m$depth_sampled_mbgl, parse = TRUE, rr.digits = 2, color = "blue", label.x = 0.2)+
  scale_y_reverse(name = latex2exp::TeX("depth$\\,$/$\\,$$\\left[$m$\\right]$"))+
    scale_x_continuous(name = latex2exp::TeX("temperature$\\,$/$\\,$$\\left[$˚C$\\right]$"))+
    theme_bw()+
  theme(
    panel.grid = element_blank()
  )

plot_T_BA3A_below_100m 
```

```{r lm-T-d-BA3A-below-100m}
# generate linear model
model_T_d_BA3A_below_100m = lm(T_C ~ depth_sampled_mbgl,
           data = BA3A_profile_below_100m)

# print model summary
summary(model_T_d_BA3A_below_100m)
```
